Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint

### 2014-11-14 egpaper_final.tex

\section{Related Works}
\subsection{Whitened Histograms of Orientations (WHO)}

\scream{Should we merge this with NZ-WHO?}
% History of Whitening and how whitening works
Recently, Hariharan \etal. introduced Whitened Histograms of Orientations (WHO)
\cite{Hariharan12}, which used statistics from natural images to whiten HOG
templates, making them more discriminative. Whitening is a common signal
processing operation for decorrelating a set of random variables
\cite{Martinsson05, Belouchrani00}. More formally, suppose that we have a
$k$-dimensional random variable $X \in \mathcal{R}^k$ with $Cov(X)=\Sigma$. By
whitening the signal, i.e., $\tilde{X}=\Sigma^{-\frac{1}{2}}(X - E[X])$, we
remove 2nd order correlation between all features. So, all elements in the
random variable $\tilde{X} = \left[\tilde{X}_1, \tilde{X}_2, \dots,
\tilde{X}_k\right]$, become more independent. In the context of HOG features, a
whitened HOG template will produce Gaussian noise when convolved with a random
natural image, but a delta-like response on an image matching the template. WHO
templates are more discriminative than standard HOG templates because they only
capture the gradients that differentiate the object from random images.
\scream{requires revision} Hariharan \etal. also interpreted the
whitening as Linear Discriminant Analysis (LDA) where all classes have different
means but have the same covariance. The whitening and LDA have the same
formulation when we make a decision boundary $w = \Sigma^{-1}(\mu_+ - \mu_0)$
where $\mu_+$ is the features from a class (LDA) or a signal to whiten
(whitening), $\mu_0$ is the negative class center (LDA) or a signal statistics
(whitening) and $\Sigma$ is the covariance matrix of classes (LDA) or
correlation within signal (whitening).

\textcolor{red}{Should we mention how our method improves on this (ie by using CG)? Since we explain in depth later, it would be okay}

@@

\textcolor{red}{Again, should we highlight the differences between these methods and ours?}

@@

We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), that are both
more discriminative than WHO, and can be generated in 70ms (compared to several
seconds for WHO). This speed up means we can generate templates on the fly,
allowing our approach to evaluate arbitrary viewpoint hypotheses. 
\scream{require revision: We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO),
an improvement over WHO which is both more discriminative than WHO and can be
computed many orders of magnitude faster than WHO (70ms as opposed to to
several seconds for WHO). This speedup allows us to generate NZ-WHO templates
on the fly, allowing us to quickly evaluate arbitrary viewpoint hypotheses.}

@@

Our second innovation is 'Non-Zero' whitening. When synthesizing detection
templates from rendered images, a common problem is how to handle the
background. If we render the model over a natural image background, gradients
in the background will be incorporated into the discriminative template.
Alternatively, we could leave the background textureless \scream{did Hariharan
do this? Nobody did this before} (see Fig.~\ref{fig:rendering}), but whitening the resulting HOG
template produces in undesirable background artifacts. These are created when
centering the template (by subtracting the mean $\mu$), where strong negative
weights are introduced in the textureless region (as seen in
Fig.~\ref{fig:whocomparison}). This could result in positive matches being
suppressed due to spurious background gradients.

>>>>

\section{Related Works}

@@

\textcolor{red}{Again, should we highlight the differences between these methods and ours? I prefer to do but wouldn't it sounds bad?}

@@

Recently, Hariharan \etal. introduced Whitened Histograms of Orientations (WHO)
\cite{Hariharan12}, which uses statistics from natural images to whiten HOG
templates, making them more discriminative. Whitening is a common signal
processing operation for decorrelating a set of random variables
\cite{Martinsson05, Belouchrani00}. More formally, suppose that we have a
$k$-dimensional random variable $X \in \mathcal{R}^k$ with $Cov(X)=\Sigma$. By
whitening the signal, i.e., $\tilde{X}=\Sigma^{-\frac{1}{2}}(X - E[X])$, we
remove 2nd order correlation between all features. So, all elements in the
random variable $\tilde{X} = \left[\tilde{X}_1, \tilde{X}_2, \dots,
\tilde{X}_k\right]$, become more independent. In the context of HOG features, a
whitened HOG template will produce Gaussian noise when convolved with a random
natural image, but a delta-like response on an image matching the template. WHO
templates are more discriminative than standard HOG templates because they only
capture the gradients that differentiate the object from random images.

Hariharan \etal. interpreted the
whitening as Linear Discriminant Analysis (LDA) where all classes have different
means but have the same covariance. The whitening and LDA have the same
formulation when we make a decision boundary $w = \Sigma^{-1}(\mu_+ - \mu_0)$
where $\mu_+$ is the features from a class (LDA) or a signal to whiten
(whitening), $\mu_0$ is the negative class center (LDA) or the signal mean
(whitening) and $\Sigma$ is the covariance matrix of classes (LDA) or
the correlation within signal (whitening).

We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), that are both
more discriminative than WHO, and can be generated several orders of magnitude
faster (in 70ms compared to several seconds for WHO). This speed up means we
can generate templates on the fly, allowing our approach to evaluate arbitrary
viewpoint hypotheses. 
%\scream{chris:We propose Non-Zero Whitened Histograms of Orientations (NZ-WHO), an improvement over WHO which is both more discriminative than WHO and can be computed many orders of magnitude faster than WHO (70ms as opposed to to several seconds for WHO). This speedup allows us to generate NZ-WHO templates on the fly, allowing us to quickly evaluate arbitrary viewpoint hypotheses.}

@@

Our second innovation is 'Non-Zero' whitening. When synthesizing detection
templates from rendered images, a common problem is how to handle the
background. If we render the model over a natural image background, gradients
in the background will be incorporated into the discriminative template.
Alternatively, we could leave the background textureless \scream{did Hariharan
do this?} (see Fig.~\ref{fig:rendering}), but whitening the resulting HOG
template produces in undesirable background artifacts. These are created when
centering the template (by subtracting the mean $\mu$), where strong negative
weights are introduced in the textureless region (as seen in
Fig.~\ref{fig:whocomparison}). This could result in positive matches being
suppressed due to spurious background gradients.
